---
title: "Basic models cascade paper"
output:
  html_document:
    df_print: paged
    code_folding: hide
    toc: true
    toc_float:
      collapsed: true
      smooth_control: true
    toc_depth: 3
    fig_caption: true
    fig_width: 4
    fig_height: 3.5
    highlight: tango
    self_contained: true
    theme: "paper"
  html_notebook:
    df_print: paged
    code_folding: hide
    toc: true
    toc_float:
      collapsed: true
      smooth_control: true
    toc_depth: 3
    fig_caption: true
    fig_width: 4
    fig_height: 3.5
    highlight: tango
    self_contained: true
    theme: "paper"
  #pdf_document: default
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(comment=NA, fig.width=4, fig.height=3.5, tidy = TRUE)
```
```{r libraries, warning=FALSE, message=FALSE, echo = FALSE}
library(deSolve)
# library(deTestSet)
library(tidyverse)
```
Basic resource and pollution models used in the cascades paper with Anne-Sophie Crepin.

## Resource

$$
\frac{dx_i}{dt} = r_i x_i \left( 1 - \frac{x_i}{K_i} \right) -
C_i\frac{(x_i)^{a_i}}{q_i^{a_i} + (x_i)^{a_i}} +
\sum_{j \neq i}D_{ij}(x_j - x_i)
$$
```{r initial_conditions}

n <- 2 ## number of cresources

## Parameters: The delta and d parameters are set here by hand to replicate ASC results.
##  Later they are set randomly on the fly.
R <- 1          # Growth rate
K <- 10 # carryng capacity
delta <- 0.05   # share of stock that diffuses to other plots
a <- 4          # maximum level after crossing the threshold
c <- 1.7        # consumption efficiency
d <- 0.02       # total diffiusion from all other plots to i
q <- 1
Dij <- matrix(c(0, delta, d, 0), nrow = 2, byrow = FALSE)

## Initial conditions:

yini <- runif(n, 8, 20)  

```

```{r, fig.align="center"}
## small dataframe
df <- tibble( x = rep(0:12))
## components of the function, with and without diffusion
growth <- function(x) {R*x*(1-(x/K))}
growth_diffusion <- function(x) {R*x*(1-(x/K)) + d*x - delta *x}
consumption <- function(x){c *((x^a) / ((q^a) + (x^a) ))}
## plot
ggplot(data = df, aes(x)) +
  stat_function(fun = growth, color = "blue", size = 0.1) +
  stat_function(fun = growth_diffusion, color = "orange", size = 0.1) +
  stat_function(fun = consumption, color = "purple", size = 0.1) +
  geom_hline(yintercept = 0, color = "gray", size = 0.1) + 
  labs(x = "x", y = "f(x)") + #xlim(0,12) + ylim(0,3) +
  theme_light(base_size = 7)

```

#### Re-scaled resource:
$$
\frac{dX_i}{d\tau} = R_i X_i \left( 1 - \frac{X_i}{K_i} \right) -
C_i\frac{(X_i)^{a_i}}{1 + (X_i)^{a_i}} +
\sum_{j \neq i}D_{ij}Q_{ij}(X_j - X_i)
$$
```{r}

#n <- 2 ## number of cresources

## Parameters: The delta and d parameters are set here by hand to replicate ASC results.
##  Later they are set randomly on the fly.
R <- 1          # Growth rate
K <- 10 # carryng capacity
delta <- 0.05   # share of stock that diffuses to other plots
a <- 4          # maximum level after crossing the threshold
c <- 2        # consumption efficiency
d <- 0.02       # total diffiusion from all other plots to i
q <- 1
Dij <- matrix(c(0, delta, d, 0), nrow = 2, byrow = FALSE)

## Initial conditions:

yini <- runif(n, 8, 20)  

```

```{r, fig.align="center"}
## small dataframe
df <- tibble( x = seq(0,10, by = 0.01))
## components of the function, with and without diffusion
growth <- function(x) {R*x*(1-(x/K))}
growth_diffusion <- function(x) {R*x*(1-(x/K)) + d*x - delta *x}
consumption <- function(x){c *((x^a) / (1 + (x^a) ))}
## plot
ggplot(data = df, aes(x)) +
  stat_function(fun = growth, color = "blue", size = 0.1) +
  stat_function(fun = growth_diffusion, color = "orange", size = 0.1) +
  stat_function(fun = consumption, color = "purple", size = 0.1) +
  geom_hline(yintercept = 0, color = "gray", size = 0.1) + 
  labs(x = "x", y = "f(x)") + #xlim(0,12) + ylim(0,3) +
  theme_light(base_size = 7)

```


### Pollution

$$\forall{i} \in\{1,...,n\}; \frac{dy_i}{dt} = u_i - b_iy_i + v_i \frac{y_{i}^{\beta_{i}}}{z_i^{\beta_i} + y_{i}^{\beta_{i}}} 
+ \sum_{j \neq i} \delta_{ij} (y_{j} - y_{i})$$

```{r warning=FALSE, message=FALSE, fig.align="center"}
## parameters on the working environment
  v = 1            # maximum rate of internal nutrient loading (top of line)
  u = 0.05           # pollution loading
  z = 1             # anoxic level
  beta = 4          # param that shapes the sharpness of the curve
  Dij = Dij         # adjacency matrix
  b = 1/2           # saturation rate (nutrient loss due to sedimentation)


## small dataframe
df <- tibble( x = rep(0:12, 3))
## components of the function, with and without diffusion
pollutant <- function(y) { v * (y^beta/(z^beta + y^beta))}  ## Original model formulated by ASC
pollutant_sedimentation <- function(y) { v * (y^beta/(z^beta + y^beta)) } ## adding the saturation parameter
management <- function(y){- u + (b*y)} ## now the diffusion term is in the management line.
management_diffusion <- function(y){ - ((0.01 * y) - (0.05 * y)) - u + (b*y)} ## now the diffusion term is in the management line.
## plot
ggplot(data = df, aes(x)) +
  #stat_function(fun = pollutant, color = "blue", linetype = 2) +
  stat_function(fun = pollutant_sedimentation, color = "orange") +
  stat_function(fun = management_diffusion, color = "purple") +
  stat_function(fun = management, color = "dodgerblue") +
  geom_hline(yintercept = 0, color = "gray") + 
  labs(x = "y", y = "f(y)") + xlim(0,3) +
  theme_light(base_size = 7)
```


Using here the modifications from Mäler et al (2003). The rescale the lake equation as:

$$\dot x(t) = a(t) - bx(t) + \frac{x^2(t)}{x^2(t) + 1}, x(0) = x_0$$

Now, notice that the anoxic level $z$ is now 1, but also $v$ dissapear, which means that it is implicitly one. All scaled values of this model are between 0-1 for $x$, and $v$ in a way is the equivalent of the carrying capacity or the maximum level of nutrients. For the Mäler model, it is bistable when parameter $1/2 < b < 3\sqrt{3/8}$ 

```{r}
## parameters on the working environment
  v = 1            # maximum rate of internal nutrient loading (top of line)
  u = 0.1           # pollution loading
  z = 1           # anoxic level
  beta = 2          # param that shapes the sharpness of the curve
  Dij = Dij         # adjacency matrix
  b = 1/2           # saturation rate (nutrient loss due to sedimentation)

## small dataframe
df <- tibble( x = seq(0,2,0.01), a = seq(0,2,0.01))
## components of the function, with and without diffusion
pollutant <- function(y) { v * (y^beta/(z^beta + y^beta))}  ## Original model formulated by ASC
pollutant_sedimentation <- function(y) { v * (y^beta/(z^beta + y^beta)) } ## adding the saturation parameter
management <- function(y){- u + (b*y)} ## now the diffusion term is in the management line.
management_diffusion <- function(y){ - ((0.01 * y) - (0.05 * y)) - u + (b*y)} ## now the diffusion term is in the management line.

## plot
ggplot(data = df, aes(x)) +
  #stat_function(fun = pollutant, color = "blue", linetype = 2) +
  stat_function(fun = pollutant, color = "orange") +
  stat_function(fun = management_diffusion, color = "purple") +
  stat_function(fun = management, color = "dodgerblue") +
  geom_hline(yintercept = 0, color = "gray") + 
  labs(x = "y", y = "f(y)") +
  theme_light(base_size = 7)
```

```{r}
## solve for a
solving_a <- function(x){
    a <- 0.48*x - ((x^2) + 1/ (x^2))
    return(a)
}

dat <- tibble(x = seq(0,2,0.01)) 
dat %>%
    mutate(b_05 = solving_a(x)) %>%
    ggplot(aes(x,b_05)) +
    geom_line()

```

