---
title: "Phase diagrams"
output:
  html_document:
    df_print: paged
    code_folding: hide
    toc: true
    toc_float:
      collapsed: true
      smooth_control: true
    toc_depth: 3
    fig_caption: true
    fig_width: 7
    fig_height: 3.5
    highlight: default
    self_contained: false
    theme: "paper"
    lib_dir: libs
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(comment=NA, fig.width=7, fig.height=3.5, tidy = TRUE)

```


```{r libraries, warning=FALSE, message=FALSE}
library(deSolve)
# library(deTestSet)
library(tidyverse)
library(patchwork)
library(network)
library(ggnetwork)
theme_set(theme_light())
```
Set up the model:

## Normalized resource model:

$$  
\frac{dY_i}{d\tau} =R_iY_i\left(1-\frac{Y_i}{K_i}\right)-C_i \frac{Y_i^{\beta_i}}{1+Y_i^{\beta_i}} + \sum_{j\not=i}D_{ij}(Q_{ij}Y_j-Y_i)
$$

Let $r$ denote the time difference between $t$ and $\tau$, $R_i=\frac{r_i}{r}$, $K_i=\frac{k_i}{q_i}$, $C_i=\frac{c_i}{rq_i}$, $Q_{ij}=\frac{d_{ji}q_j}{d_{ij}q_i}$, and $D_{ij}=\frac{d_{ij}}{r}$. 

```{r}
# This is where we define your event function
# Add this directly above your call to ode()
posfun <- function(t, y, parms){
  with(as.list(y), {
    y[which(y<0)] <- 0
    return(y)
  })
}


## model
resource_norm <- function(t, y, params){
  with(as.list(c(y, params)), {
    ## Normalization steps:
    Y = y/r0
    R = r/r0
    K0 = K/q
    C = c/(r0*q)
    Qij_parts <- d_ij * q
    Qij = t(Qij_parts) / Qij_parts ## the same as lower triangle / upper triangle and viceversa
    ## the next line corrects the case when Zij is NaN induced by division by zero. In that case, Zij should be a matrix full of zeroes = no connections.
    if(all(is.nan(Qij))) Qij = Qij_parts
    
    # set diag to zero
    diag(Qij) <- 0
    D_ij = d_ij / r0
    
    ## System with normalized parameters
    resource <- R * Y * (1 - (Y/K0)) - C * (Y^b / (1 + Y^b))
    outflow <-  (A_ij * D_ij * Qij) %*% Y
    inflow <-  t(A_ij * D_ij) %*% Y 
    # Qij is not multiplied by D_ij by suggestion of ASC.

    dX <- resource + (inflow - outflow)
    
    return(list(c(dX)))
  })
}

## set up flux matrix
n <- 1        # number of systems
#delta_ij <- matrix(runif(n^2, min = 0.02, max = 0.05), ncol = n)
d_ij <- matrix(rep(0,n^2), ncol = n) # turn off diffusion terms
diag(d_ij) <- 0
A_ij <- matrix(rbinom(n^2, 1, prob = 0.3), n, n)
diag(A_ij) <- 0


params <- list(
    r0 = rep(1,n),                # scaling parameter
    r = rep(1.5, n),              # intrinsic growth rate
    K = rep(10, n),               # Carrying capacity
    c = rep(2, n),                # max predation rate
    q = 1, #runif(n, min = 2, max = 8),        # threshold
    b = 4 ,                       # sharpness of the shift
    d_ij = d_ij,                  # matrix of diffusion terms
    A_ij = A_ij                   # adjacency matrix
)

## set up time steps
times <- seq(from = 0, to = 100, by = 0.01)

## initial conditions
yini <- runif(n, 0.5, 5)

## run the model
out <- ode(
  y = yini, times = times,  func = resource_norm, parms = params,
  method = "bdf", ## see help("ode") for more methods
  events=list(func = posfun, time = times)
)


df <- out %>% as_tibble() %>%
    gather(key = "species", value = "population", 2:last_col())

df %>%
    ggplot(aes(x=time, y=population)) +
    geom_line(aes(color = species), size = 0.25, show.legend = T) +
    scale_y_continuous(limits = c(0,NA))
```

```{r}
## make up some initial values
xini <- seq(0.25,5, 0.5)
qini <- seq(0.25,5, 0.5)

ini_grid <- expand.grid(xini, qini) %>%
    rename(xini = Var1, qini = Var2)
```

```{r}
## making simulations shorter to save some time
times <- seq(from = 0, to = 10, by = 0.01)
library(tictoc)

results <- list()
tic()
results <- pmap(
    list(
        x = ini_grid$xini,
        y = ini_grid$qini),
    .f = function(x,y){
        params$q <- y
        yini <- x
        out <- ode(
            y = yini, times = times,  func = resource_norm, parms = params,
            method = "bdf", ## see help("ode") for more methods
            events=list(func = posfun, time = times))
        return(out)
    }
)
toc()
```

```{r}
results <- results %>%
  map(
    function(x) {
      x <- x %>% 
        as_tibble() %>% 
        rename(sp = `1`) %>%
        mutate(sp_lag = tsibble::difference(sp)) # this is the derivative
      return(x)
    }
  )

```

```{r}
results <- results %>% 
  map2(., ini_grid$xini, .f = function(x,y) {
    x$xini <- y
    return(x)
  })

results <- results %>% 
  map2(., ini_grid$qini, .f = function(x,y) {
    x$qini <- y
    return(x)
  })
```

```{r}
results %>%
  map(., function(x) x %>% map_df(., as.numeric)) %>% 
  bind_rows() %>%
  mutate(id = paste(xini, qini, sep = "_")) %>%
  ggplot(aes(sp, sp_lag, group = id)) +
  geom_hline(yintercept = 0, color = "black", size = 0.1) +
  geom_line(aes(color = as_factor(qini)), size = 0.2, show.legend = TRUE) + 
  labs(x = "x", y = "dx", title = "Sensitivity to innitial conditions and threshold q" ,
       subtitle = "qini are different parameter values of q") +
  facet_wrap(~xini)
```

```{r}
results %>%
  map(., function(x) x %>% map_df(., as.numeric)) %>% 
  bind_rows() |> 
  filter(time == 0.01 | time == 1) 
```



### 2 species

```{r}
ini_grid <- expand.grid(1:5, 1:5, 1:5) 

## set up flux matrix
n <- 2        # number of systems
d_ij <- matrix(runif(n^2, min = 0.02, max = 0.05), ncol = n)
#d_ij <- matrix(rep(0,n^2), ncol = n) # turn off difussion terms
diag(d_ij) <- 0
A_ij <- matrix(rbinom(n^2, 1, prob = 0.3), n, n)
diag(A_ij) <- 0


params <- list(
    r0 = rep(1,n),                # scaling parameter
    r = rep(1.5, n),              # intrinsic growth rate
    K = rep(10, n),               # Carrying capacity
    c = rep(2, n),                # max predation rate
    q = 4, #runif(n, min = 2, max = 8),        # threshold
    b = 4 ,                       # sharpness of the shift
    d_ij = d_ij,                  # matrix of diffusion terms
    A_ij = A_ij                   # adjacency matrix
)

## set up time steps
times <- seq(from = 0, to = 10, by = 0.01)

## initial conditions
# yini <- runif(n, 0.5, 5)
```

```{r}
results <- list()
tic()
results <- pmap(
    ini_grid,
    .f = function(Var1,Var2,Var3){
        yini <- c(Var1,Var2)
        params$q <- Var3
        out <- ode(
            y = yini, times = times,  func = resource_norm, parms = params,
            method = "bdf", ## see help("ode") for more methods
            events=list(func = posfun, time = times))
        return(out)
    }
)
toc()
```

```{r}
results <- results %>%
  map(
    function(x) {
      x <- x %>% 
        as_tibble() %>% 
        rename(sp1 = `1`, sp2 = `2`)
      return(x)
    }
  )

```
```{r}
results <- results %>% 
  map2(., 1:nrow(ini_grid), .f = function(x,y) {
    x$group <- y
    return(x)
  })
```

```{r}

results <- results %>% 
  map2(., ini_grid$Var3, .f = function(x,y) {
    x$qini <- y
    return(x)
  })
```


```{r}
results %>%
  map(., function(x) x %>% map_df(., as.numeric)) %>% 
  bind_rows() %>%
  #mutate(id = paste(xini, qini, sep = "_")) %>%
  ggplot(aes(sp1, sp2, group = group)) +
  geom_line( #aes(color = as_factor(qini)), 
    linetype = 2, size = 0.1, show.legend = TRUE) +
  facet_wrap(~qini) +
  coord_equal()
```


TODO: Extract fields using a lag per variable (doesn't have to be lag 1), and then sample the dataset so you have less arrows to draw.



## leftovers

### phase diagrams

The only way to change $X$ is through $x_i$, same with $U$ and $u_i$. To better
understand the param space where the bifurcation occurs, I need first to model
individual systems. So back to the basic model. **not true, ASC pointed out I
can use sp1 vs sp2**

```{r}
# ## set up flux matrix
# n <- 10
# #delta_ij <- matrix(runif(n^2, min = 0.02, max = 0.05), ncol = n)
# delta_ij <- matrix(rep(0,n^2), ncol = n) # turn off difussion terms
# diag(delta_ij) <- 0
# A_ij <- matrix(0, n, n) # no connections
# diag(A_ij) <- 0
# ## Parameters: use rep() if the value is the same, or runif() ir meant to be different across systems
# params <- list(
#     u = seq.int(0.5,10,length.out = n),              # pollution load from humans
#     s = rep(0.7, n),              # internal loss rate (sedimentation)
#     v = rep(5, n),               # max level of internal nutrient release
#     z = rep(4, n),                # threshold, same for all
#     alpha = 4 ,                   # sharpness of the shift
#     delta_ij = delta_ij,          # matrix of difussion terms
#     A_ij = A_ij                   # adjacency matrix
# )
# 
# ## set up time steps
# times <- seq(from = 0, to = 100, by = 0.01)
# 
# ## initial conditions: more systematic
# yini <- seq.int(from =0.5,to=5, length.out = n)
# 
# ## run the model
# print(system.time(
#     out <- ode(
#       y = yini, times = times,  func = pollution, parms = params,
#       method = "bdf" , ## see help("ode") for more methods
#       events=list(func = posfun, time = times)
#     )
# ))
# 
# df <- out %>% as_tibble() %>%
#     gather(key = "species", value = "population", 2:last_col())
# 
# df %>%
#     ggplot(aes(x=time, y=population)) +
#     geom_line(aes(color = species), size = 0.25, show.legend = T) + 
#   #scale_x_log10() +
#     theme_light()
```

What are sensible values?

```{r}
# ## small dataframe
# df <- tibble( x = rep(0:12, 3))
# ## components of the function, with and without diffusion
# pollutant <- function(x) {  v * (x^alpha/(z^alpha + x^alpha))}  ## Original model formulated by ASC
# management <- function(x){ -u + s*x } ## now the diffusion term is in the management line.
# management_diffusion <- function(x){ - ((0.01 * x) - (0.05 * x)) - u + (s*x)} ## now the diffusion term is in the management line.
# 
# #parameter selection:
# #params <- list(
#     u = 0.5                # pollution load from humans
#     s = 0.7              # internal loss rate (sedimentation)
#     v = 5               # max level of internal nutrient release
#     z = 4                # threshold, same for all
#     alpha = 4            # sharpness of the shift
#     delta_ij = delta_ij  # matrix of difussion terms
#     A_ij = A_ij          # adjacency matrix
# # )
# ## plot
# ggplot(data = df, aes(x)) +
#   #stat_function(fun = pollutant, color = "blue", linetype = 2) +
#   stat_function(fun = pollutant, color = "orange") +
#   stat_function(fun = management_diffusion, color = "purple") +
#   stat_function(fun = management, color = "dodgerblue") +
#   geom_hline(yintercept = 0, color = "gray") + 
#   labs(x = "y", y = "f(y)") +
#   theme_light()
# ```
# 
# ```{r}
# ## phase diagram attempt:
# 
# u <- seq.int(from =0.5,to=10, length.out = 10)
# x_0 <- seq.int(from =0.5,to=10, length.out = 10)
# 
# df_sims <- expand.grid(u, x_0) %>%
#   rename(u = Var1, x_0 = Var2)
# 
# ## for the phase diagram I only need short time snips: set up time steps
# times <- seq(from = 0, to = 2, by = 0.01)
# 
# sims <- list()
# 
# sims <- map2(
#   df_sims$u, df_sims$x_0,
#   function(x,y){
#     params$u <- x
#     yini <- y
#     out <- ode(
#       y = yini, times = times,  func = pollution, parms = params,
#       method = "bdf" , ## see help("ode") for more methods
#       events=list(func = posfun, time = times))
#     return(out)
#   }
# )

```

```{r}
# l <- length(sims)
# sims <- sims %>%
#   map(function(x)  {x %>% as_tibble() %>%
#     gather(key = "species", value = "population", 2:last_col())}) %>%
#   map2(., df_sims$u, function(x,y) {
#     x$u <- y
#     return(x)}) %>%
#   map2(., .y=seq(from =1, to =l,by =1), .f = function(x,y) {
#     x$id <- y 
#     return(x)}) %>%
#   bind_rows()
#   
```

```{r}
# sims %>%
#   map_df(as.numeric) %>%
#   group_by(id) %>%
#   mutate(pop_lag = lag(population)) %>%
#   ggplot(aes(population, pop_lag)) +
#   geom_path(aes(color = u, group = id), size = 0.5) +
#   theme_light()

```






### more left overs

### phase diagrams

The only way to change $X$ is through $x_i$, same with $U$ and $u_i$. To better understand the param space where the bifurcation occurs, I need first to model individual systems. So back to the basic model.

```{r}
## set up flux matrix
# n <- 10
# #delta_ij <- matrix(runif(n^2, min = 0.02, max = 0.05), ncol = n)
# delta_ij <- matrix(rep(0,n^2), ncol = n) # turn off difussion terms
# diag(delta_ij) <- 0
# A_ij <- matrix(0, n, n) # no connections
# diag(A_ij) <- 0
# ## Parameters: use rep() if the value is the same, or runif() ir meant to be different across systems
# params <- list(
#     u = seq.int(0.5,10,length.out = n),              # pollution load from humans
#     s = rep(0.7, n),              # internal loss rate (sedimentation)
#     v = rep(5, n),               # max level of internal nutrient release
#     z = rep(4, n),                # threshold, same for all
#     alpha = 4 ,                   # sharpness of the shift
#     delta_ij = delta_ij,          # matrix of difussion terms
#     A_ij = A_ij                   # adjacency matrix
# )
# 
# ## set up time steps
# times <- seq(from = 0, to = 100, by = 0.01)
# 
# ## initial conditions: more systematic
# yini <- seq.int(from =0.5,to=5, length.out = n)
# 
# ## run the model
# print(system.time(
#     out <- ode(
#       y = yini, times = times,  func = pollution, parms = params,
#       method = "bdf" , ## see help("ode") for more methods
#       events=list(func = posfun, time = times)
#     )
# ))
# 
# df <- out %>% as_tibble() %>%
#     gather(key = "species", value = "population", 2:last_col())
# 
# df %>%
#     ggplot(aes(x=time, y=population)) +
#     geom_line(aes(color = species), size = 0.25, show.legend = T) + 
#   scale_x_log10() +
#     theme_light()
```
What are sensible values?

```{r}
# ## small dataframe
# df <- tibble( x = rep(0:12, 3))
# ## components of the function, with and without diffusion
# pollutant <- function(x) {  v * (x^alpha/(z^alpha + x^alpha))}  ## Original model formulated by ASC
# management <- function(x){ -u + s*x } ## now the diffusion term is in the management line.
# management_diffusion <- function(x){ - ((0.01 * x) - (0.05 * x)) - u + (s*x)} ## now the diffusion term is in the management line.
# 
# #parameter selection:
# #params <- list(
#     u = 0.5                # pollution load from humans
#     s = 0.7              # internal loss rate (sedimentation)
#     v = 5               # max level of internal nutrient release
#     z = 4                # threshold, same for all
#     alpha = 4            # sharpness of the shift
#     delta_ij = delta_ij  # matrix of difussion terms
#     A_ij = A_ij          # adjacency matrix
# # )
# ## plot
# ggplot(data = df, aes(x)) +
#   #stat_function(fun = pollutant, color = "blue", linetype = 2) +
#   stat_function(fun = pollutant, color = "orange") +
#   stat_function(fun = management_diffusion, color = "purple") +
#   stat_function(fun = management, color = "dodgerblue") +
#   geom_hline(yintercept = 0, color = "gray") + 
#   labs(x = "y", y = "f(y)") +
#   theme_light()
```

```{r}
## phase diagram attempt:
# 
# u <- seq.int(from =0.5,to=10, length.out = 10)
# x_0 <- seq.int(from =0.5,to=10, length.out = 10)
# 
# df_sims <- expand.grid(u, x_0) %>%
#   rename(u = Var1, x_0 = Var2)
# 
# sims <- list()
# 
# sims <- map2(
#   df_sims$u, df_sims$x_0,
#   function(x,y){
#     params$u <- x
#     yini <- y
#     out <- ode(
#       y = yini, times = times,  func = pollution, parms = params,
#       method = "bdf" , ## see help("ode") for more methods
#       events=list(func = posfun, time = times))
#     return(out)
#   }
# )

```

```{r}
# l <- length(sims)
# sims <- sims %>%
#   map(function(x)  {x %>% as_tibble() %>%
#     gather(key = "species", value = "population", 2:last_col())}) %>%
#   map2(., df_sims$u, function(x,y) {
#     x$u <- y
#     return(x)}) %>%
#   map2(., .y=seq(from =1, to =l,by =1), .f = function(x,y) {
#     x$id <- y 
#     return(x)}) %>%
#   bind_rows()
#   
```

```{r}
# sims %>%
#   map_df(as.numeric) %>%
#   group_by(id) %>%
#   mutate(pop_lag = lag(population)) %>%
#   ggplot(aes(population, pop_lag)) +
#   geom_path(aes(color = u, group = id), size = 0.5) +
#   theme_light()

```

