---
title: "Resource model"
output: html_notebook
---

```{r libraries, warning=FALSE, message=FALSE}
library(deSolve)
# library(deTestSet)
library(tidyverse)
theme_set(theme_light())
```
Updated version (J200925) with the current notation of the paper

$$\forall{i} \in\{1,...,n\}; \frac{dy_i}{dt} = r_iy_i\left(1-\frac{y_i}{k_i}\right) 
-c_i \frac{y_i^{\beta_i}}{q_i^{\beta_i}+y_i^{\beta_i}} 
+ \sum_{j\not=i}(d_{ij}y_j-d_{ji}y_i). $$

Notice that the second term can be factorized and extract $d_{ij}$ in matrix form and simply multiply by the difference $(y_i - y_j)$.

```{r}

# This is where we define your event function
# Add this directly above your call to ode()
posfun <- function(t, y, parms){
  with(as.list(y), {
    y[which(y<0)] <- 0
    return(y)
  })
}

## Resource model:

resource <- function(t, y, params){
  with(as.list(c(y, params)), {
    growth <- r * y * (1- y/K)  ## This should recover Lodka-Volterra
    consumption <- c *((y^b) / ((q^b) + (y^b) ))
        
    immigration <-  t(A_ij *d_ij) %*% y
    emmigration <-  (A_ij * d_ij) %*% y
    
    dx <- growth - consumption + (immigration - emmigration)

    return(list(c(dx)))
  })
}

## set up flux matrix
n <- 4
#delta_ij <- matrix(runif(n^2, min = 0.02, max = 0.05), ncol = n)
d_ij <- matrix(rep(0,n^2), ncol = n) # turn off difussion terms
diag(d_ij) <- 0
A_ij <- matrix(rbinom(n^2, 1, prob = 0.3), n, n)
diag(A_ij) <- 0
## Parameters: use rep() if the value is the same, or runif() ir meant to be different across systems
params <- list(
    r = rep(1.5, n),              # intrinsic growth rate
    K = rep(10, n),               # Carrying capacity
    c = rep(2, n),                # max predation rate
    q = runif(n, min = 2, max = 8),        # threshold
    b = 4 ,                       # sharpness of the shift
    d_ij = d_ij,                  # matrix of difussion terms
    A_ij = A_ij                   # adjacency matrix
)

## set up time steps
times <- seq(from = 0, to = 100, by = 0.01)

## initial conditions
yini <- runif(n, 5, 20)

## run the model
#print(system.time(
    out <- ode(
      y = yini, times = times,  func = resource, parms = params,
      #method = "bdf" , ## see help("ode") for more methods
      events=list(func = posfun, time = times)
    )
#))

```
Model with no connections:

```{r warning = FALSE, message = FALSE}
df <- out %>% as_tibble() %>%
    gather(key = "species", value = "population", 2:last_col())

df %>%
    ggplot(aes(x=time, y=population)) +
    geom_line(aes(color = species), size = 0.25, show.legend = T) + 
    theme_light()
```

Connected model: 

```{r warning = FALSE, message = FALSE}
## set up flux matrix
d_ij <- matrix(runif(n^2, min = 0.02, max = 0.05), ncol = n)
diag(d_ij) <- 0

## Only change the delta:
params$d_ij <-  d_ij           # adjacency matrix of difussion terms

## run the model
out <- ode(
  y = yini, times = times,  func = resource, parms = params,
  method = "bdf", ## see help("ode") for more methods
  events=list(func = posfun, time = times)
)


df <- out %>% as_tibble() %>%
    gather(key = "species", value = "population", 2:last_col())

df %>%
    ggplot(aes(x=time, y=population)) +
    geom_line(aes(color = species), size = 0.25, show.legend = T) + 
    theme_light()
```



## Normalized model:

$$  
\frac{dY_i}{d\tau} =R_iY_i\left(1-\frac{Y_i}{K_i}\right)-C_i \frac{Y_i^{\beta_i}}{1+Y_i^{\beta_i}} + \sum_{j\not=i}D_{ij}(Q_{ij}Y_j-Y_i)
$$

Let $r$ denote the time difference between $t$ and $\tau$, $R_i=\frac{r_i}{r}$, $K_i=\frac{k_i}{q_i}$, $C_i=\frac{c_i}{rq_i}$, $Q_{ij}=\frac{d_{ji}q_j}{d_{ij}q_i}$, and $D_{ij}=\frac{d_{ij}}{r}$. 

```{r warning = FALSE, message = FALSE}
## Pollution model
resource_norm <- function(t, y, params){
  with(as.list(c(y, params)), {
    ## Normalization steps:
    Y = y/r0
    R = r/r0
    K0 = K/q
    C = c/(r0*q)
    Qij_parts <- d_ij * q
    Qij = t(Qij_parts) / Qij_parts ## the same as lower triangle / upper triangle and viceversa
    ## the next line corrects the case when Zij is NaN induced by division by zero. In that case, Zij should be a matrix full of zeroes = no connections.
    if(all(is.nan(Qij))) Qij = Qij_parts
    
    # set diag to zero
    diag(Qij) <- 0
    D_ij = d_ij / r0
    
    ## System with normalized parameters
    resource <- R * Y * (1 - (Y/K0)) - C * (Y^b / (1 + Y^b))
    outflow <-  (A_ij * D_ij * Qij) %*% Y
    inflow <-  t(A_ij * D_ij) %*% Y 
    # Qij is not multiplied by D_ij by suggestion of ASC.

    dX <- resource + (inflow - outflow)
    
    return(list(c(dX)))
  })
}

params <- list(
    r0 = rep(1,n),                # scaling parameter
    r = rep(1.5, n),              # intrinsic growth rate
    K = rep(10, n),               # Carrying capacity
    c = rep(2, n),                # max predation rate
    q = runif(n, min = 2, max = 8),        # threshold
    b = 4 ,                       # sharpness of the shift
    d_ij = d_ij,                  # matrix of diffusion terms
    A_ij = A_ij                   # adjacency matrix
)

## run the model
out <- ode(
  y = yini, times = times,  func = resource_norm, parms = params,
  method = "bdf", ## see help("ode") for more methods
  events=list(func = posfun, time = times)
)


df <- out %>% as_tibble() %>%
    gather(key = "species", value = "population", 2:last_col())

df %>%
    ggplot(aes(x=time, y=population)) +
    geom_line(aes(color = species), size = 0.25, show.legend = T) + 
    theme_light()

```

## To-do:

1. Systems without connections: Now with a working model, update and change some parameters.
```{r warning=FALSE, message=FALSE}
# delta_ij # old delta_ij used in previous runs
## Turn-off the connection
#params$delta_ij <- matrix(rep(0, n^2), n, n)
params$Aij <- matrix(rep(0, n^2), n, n)

## run the model
out <- ode(
  y = yini, times = times,  func = resource_norm, parms = params,
  #method = "bdf" ## see help("ode") for more methods
  events=list(func = posfun, time = times)
)


out %>% as_tibble() %>%
    gather(key = "system", value = "pollutant", 2:last_col()) %>%
    ggplot(aes(x=time, y=pollutant)) +
    geom_line(aes(color = system), size = 0.25, show.legend = T) + 
    theme_light()
```



2. Systems with connections (max 10)

I first update N to 10 and create initial conditions for all systems. For comparison I run a model without connection, and then turn on difussion. 

```{r warning=FALSE, message=FALSE}
## I need to update all parameters then:
## set up flux matrix
n <- 10
#delta_ij <- matrix(runif(n^2, min = 0.02, max = 0.05), ncol = n)
d_ij <- matrix(rep(0,n^2), ncol = n) # turn off difussion
diag(d_ij) <- 0
A_ij <- matrix(rbinom(n^2, 1, prob = 0.3), n, n)
diag(A_ij) <- 0
## Parameters: use rep() if the value is the same, or runif() ir meant to be different across systems
params <- list(
    r0 = rep(1,n),                # scaling parameter
    r = rep(1.5, n),              # intrinsic growth rate
    K = rep(10, n),               # Carrying capacity
    c = rep(2, n),                # max predation rate
    q = runif(n, min = 2, max = 8),        # threshold
    b = 4 ,                       # sharpness of the shift
    d_ij = d_ij,                  # matrix of diffusion terms
    A_ij = A_ij                   # adjacency matrix
)

## initial conditions
yini <- runif(n, 5, 20)
## run the model
out <- ode(
  y = yini, times = times,  func = resource_norm, parms = params,
  #method = "bdf" ## see help("ode") for more methods
  events=list(func = posfun, time = times)
)

## Visualize
out %>% as_tibble() %>%
    gather(key = "system", value = "pollutant", 2:last_col()) %>%
    ggplot(aes(x=time, y=pollutant)) +
    geom_line(aes(color = system), size = 0.25, show.legend = T) + 
    theme_light()
```

Note that currently the $d_{ij}$ matrix has low diffusion coefficients (2-5%) but all lakes are connected to each other. To better control structure, one should multiply the difussion coefficents to a [0,1] adjacency matrix, so one gets control of structure on a term by itself.

```{r warning = FALSE, message=FALSE}
## get a erdos-renyi random graph
Aij <- igraph::sample_gnp(n, 0.2, TRUE) %>% 
  igraph::as_adjacency_matrix(graph = .) %>% 
  as.matrix()

## activate conections
d_ij <- matrix(runif(n^2, min = 0.02, max = 0.05), ncol = n)
diag(d_ij) <- 0
params$d_ij <- d_ij
params$A_ij <- A_ij

## run the model
out <- ode(
  y = yini, times = times,  func = resource_norm, parms = params,
  #method = "bdf" ## see help("ode") for more methods
  events=list(func = posfun, time = times)
)

# Visualize
out %>% as_tibble() %>%
    gather(key = "system", value = "pollutant", 2:last_col()) %>%
    ggplot(aes(x=time, y=pollutant)) +
    geom_line(aes(color = system), size = 0.25, show.legend = T) + 
    theme_light()
```

```{r}
## plot the network
library(network)
library(ggnetwork)
net <- network::network(A_ij, directed = T)

p1 <- ggnetwork(net) %>%
  ggplot() + 
  geom_edges(
    aes(x = x,y = y,xend = xend,yend = yend), color = "gray50", size = 0.1,
    arrow = arrow(length = unit(0.15, "lines"), type = "closed"),
    curvature = 0.1) +
  geom_nodes(aes(x =x,y = y, color = as.factor(vertex.names)), size = 2, show.legend = FALSE ) +
  theme_void(base_size = 4)
  
p1
```



3. Scenario when Q_ij > 1 vs. Q_ij < 1

Because $Q_{ij}=\frac{d_{ji}z_j}{d_{ij}z_i}$ and $q_i = q_j$, the only way to control $Q$ in a way that it becomes >1 or <1 depends on $d_{i,j}$. The matrix needs to be created with the upper triangle > lower triangle or viceversa.

```{r warning = FALSE, message=FALSE}
## This makes sure that Z_ij > 1
d_ij <- matrix(0, ncol = n, nrow = n)
d_ij[upper.tri(d_ij) ] <-  0.02
d_ij[lower.tri(d_ij) ]<- 0.05

params$d_ij <- d_ij

## run the model
out <- ode(
  y = yini, times = times,  func = resource_norm, parms = params,
  #method = "bdf" ## see help("ode") for more methods
  events=list(func = posfun, time = times)
)

## This makes sure that Z_ij > 1
d_ij <- matrix(0, ncol = n, nrow = n)
d_ij[upper.tri(d_ij) ] <-  0.05
d_ij[lower.tri(d_ij) ]<- 0.02

params$d_ij <- d_ij

## run the model
out2 <- ode(
  y = yini, times = times,  func = resource_norm, parms = params,
  #method = "bdf" ## see help("ode") for more methods
  events=list(func = posfun, time = times)
)

out <- out %>% as_tibble() %>%
  gather(key = "system", value = "pollutant", 2:last_col()) %>%
  add_column(scenario = "Z > 1") %>%
  mutate(time = as.numeric(time), pollutant = as.numeric(pollutant))
  
out2 <- out2 %>% as_tibble() %>%
  gather(key = "system", value = "pollutant", 2:last_col()) %>%
  add_column(scenario = "Z < 1") %>% 
  mutate(time = as.numeric(time), pollutant = as.numeric(pollutant))

# Visualize
p2 <- full_join(out,out2) %>% 
    ggplot(aes(x=time, y=pollutant)) +
    geom_line(aes(color = system), size = 0.25, show.legend = FALSE) + 
  facet_wrap(~scenario) +
    theme_light(base_size = 6)
```

```{r}
library(patchwork)

p2 + p1 + plot_layout(width = c(2,1)) 

# ggsave("figure_Z_comparison.png", width = 5, height = 2, units = "in", dpi = 800)
```






However, not all $Q_{ij}$ are >1 or < 1. 

```{r}
qmat <- with(params, d_ij * q)
(t(qmat)/qmat) > 1
```

The reason for that is that it depends also of $q_i$ on the original equation. The only way to control for that is to constructing $q$ by hand, or making it the same for all systems.

4. Scenario when $D_ij > 1$ versus $D_{ij} < 1$

```{r warning = FALSE, message=FALSE}
## reset d_ij
d_ij <- matrix(runif(n^2, min = 0.02, max = 0.05), ncol = n)
diag(d_ij) <- 0
params$d_ij <- d_ij

## By setting s to > min value of delta, the matrix is >=1 always
params$s <- 0.015

## run the model
out <- ode(
  y = yini, times = times,  func = resource_norm, parms = params,
  method = "bdf" ,## see help("ode") for more methods
  events=list(func = posfun, time = times)
)

## This makes sure that Z_ij > 1
## By setting s to > max value of delta, the matrix is <= 1 always
params$s <- 0.07

## run the model
out2 <- ode(
  y = yini, times = times,  func = resource_norm, parms = params,
  method = "bdf", ## see help("ode") for more methods
  events=list(func = posfun, time = times)
)

out <- out %>% as_tibble() %>%
  gather(key = "system", value = "pollutant", 2:last_col()) %>%
  add_column(scenario = "Delta_{ij} > 1") %>%
  mutate(time = as.numeric(time), pollutant = as.numeric(pollutant))
  
out2 <- out2 %>% as_tibble() %>%
  gather(key = "system", value = "pollutant", 2:last_col()) %>%
  add_column(scenario = "Delta_{ij} < 1") %>% 
  mutate(time = as.numeric(time), pollutant = as.numeric(pollutant))

# Visualize
full_join(out,out2) %>% 
    ggplot(aes(x=time, y=pollutant)) +
    geom_line(aes(color = system), size = 0.25, show.legend = T) + 
  facet_wrap(~scenario) +
    theme_light()
```

5. Test for structural stability before running simulations
6. ~~Repeat for resource system. [done]~~
7. Set up controlling experiments

Notes after meeting with ASC: 

* ~~fix the $A_{ij}$, the model was not updated [done]~~
* ~~extract results and create a graph with scenarios when Z and Delta are >1<~~
* phase diagrams X vs U?

### phase diagrams

The only way to change $X$ is through $x_i$, same with $U$ and $u_i$. To better understand the param space where the bifurcation occurs, I need first to model individual systems. So back to the basic model.

```{r}
## set up flux matrix
n <- 10
#delta_ij <- matrix(runif(n^2, min = 0.02, max = 0.05), ncol = n)
delta_ij <- matrix(rep(0,n^2), ncol = n) # turn off difussion terms
diag(delta_ij) <- 0
A_ij <- matrix(0, n, n) # no connections
diag(A_ij) <- 0
## Parameters: use rep() if the value is the same, or runif() ir meant to be different across systems
params <- list(
    u = seq.int(0.5,10,length.out = n),              # pollution load from humans
    s = rep(0.7, n),              # internal loss rate (sedimentation)
    v = rep(5, n),               # max level of internal nutrient release
    z = rep(4, n),                # threshold, same for all
    alpha = 4 ,                   # sharpness of the shift
    delta_ij = delta_ij,          # matrix of difussion terms
    A_ij = A_ij                   # adjacency matrix
)

## set up time steps
times <- seq(from = 0, to = 100, by = 0.01)

## initial conditions: more systematic
yini <- seq.int(from =0.5,to=5, length.out = n)

## run the model
print(system.time(
    out <- ode(
      y = yini, times = times,  func = pollution, parms = params,
      method = "bdf" , ## see help("ode") for more methods
      events=list(func = posfun, time = times)
    )
))

df <- out %>% as_tibble() %>%
    gather(key = "species", value = "population", 2:last_col())

df %>%
    ggplot(aes(x=time, y=population)) +
    geom_line(aes(color = species), size = 0.25, show.legend = T) + 
  scale_x_log10() +
    theme_light()
```
What are sensible values?

```{r}
## small dataframe
df <- tibble( x = rep(0:12, 3))
## components of the function, with and without diffusion
pollutant <- function(x) {  v * (x^alpha/(z^alpha + x^alpha))}  ## Original model formulated by ASC
management <- function(x){ -u + s*x } ## now the diffusion term is in the management line.
management_diffusion <- function(x){ - ((0.01 * x) - (0.05 * x)) - u + (s*x)} ## now the diffusion term is in the management line.

#parameter selection:
#params <- list(
    u = 0.5                # pollution load from humans
    s = 0.7              # internal loss rate (sedimentation)
    v = 5               # max level of internal nutrient release
    z = 4                # threshold, same for all
    alpha = 4            # sharpness of the shift
    delta_ij = delta_ij  # matrix of difussion terms
    A_ij = A_ij          # adjacency matrix
# )
## plot
ggplot(data = df, aes(x)) +
  #stat_function(fun = pollutant, color = "blue", linetype = 2) +
  stat_function(fun = pollutant, color = "orange") +
  stat_function(fun = management_diffusion, color = "purple") +
  stat_function(fun = management, color = "dodgerblue") +
  geom_hline(yintercept = 0, color = "gray") + 
  labs(x = "y", y = "f(y)") +
  theme_light()
```

```{r}
## phase diagram attempt:

u <- seq.int(from =0.5,to=10, length.out = 10)
x_0 <- seq.int(from =0.5,to=10, length.out = 10)

df_sims <- expand.grid(u, x_0) %>%
  rename(u = Var1, x_0 = Var2)

sims <- list()

sims <- map2(
  df_sims$u, df_sims$x_0,
  function(x,y){
    params$u <- x
    yini <- y
    out <- ode(
      y = yini, times = times,  func = pollution, parms = params,
      method = "bdf" , ## see help("ode") for more methods
      events=list(func = posfun, time = times))
    return(out)
  }
)

```

```{r}
l <- length(sims)
sims <- sims %>%
  map(function(x)  {x %>% as_tibble() %>%
    gather(key = "species", value = "population", 2:last_col())}) %>%
  map2(., df_sims$u, function(x,y) {
    x$u <- y
    return(x)}) %>%
  map2(., .y=seq(from =1, to =l,by =1), .f = function(x,y) {
    x$id <- y 
    return(x)}) %>%
  bind_rows()
  
```

```{r}
sims %>%
  map_df(as.numeric) %>%
  group_by(id) %>%
  mutate(pop_lag = lag(population)) %>%
  ggplot(aes(population, pop_lag)) +
  geom_path(aes(color = u, group = id), size = 0.5) +
  theme_light()

```

