---
title: "Resource model"
output:
  html_document:
    df_print: paged
    code_folding: hide
    toc: true
    toc_float:
      collapsed: true
      smooth_control: true
    toc_depth: 3
    fig_caption: true
    fig_width: 7
    fig_height: 3.5
    highlight: default
    self_contained: false
    theme: "paper"
    lib_dir: libs
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(comment=NA, fig.width=7, fig.height=3.5, tidy = TRUE)

```



```{r libraries, warning=FALSE, message=FALSE}
library(deSolve)
# library(deTestSet)
library(tidyverse)
library(patchwork)
library(network)
library(ggnetwork)
theme_set(theme_light())
```
Updated version (J200925) with the current notation of the paper

$$\forall{i} \in\{1,...,n\}; \frac{dy_i}{dt} = r_iy_i\left(1-\frac{y_i}{k_i}\right) 
-c_i \frac{y_i^{\beta_i}}{q_i^{\beta_i}+y_i^{\beta_i}} 
+ \sum_{j\not=i}(d_{ij}y_j-d_{ji}y_i). $$

Notice that the second term can be factorized and extract $d_{ij}$ in matrix form and simply multiply by the difference $(y_i - y_j)$.

```{r}

# This is where we define your event function
# Add this directly above your call to ode()
posfun <- function(t, y, parms){
  with(as.list(y), {
    y[which(y<0)] <- 0
    return(y)
  })
}

## Resource model:

resource <- function(t, y, params){
  with(as.list(c(y, params)), {
    growth <- r * y * (1- y/K)  ## This should recover Lodka-Volterra
    consumption <- c *((y^b) / ((q^b) + (y^b) ))
        
    immigration <-  t(A_ij *d_ij) %*% y
    emmigration <-  (A_ij * d_ij) %*% y
    
    dx <- growth - consumption + (immigration - emmigration)

    return(list(c(dx)))
  })
}

## set up flux matrix
n <- 4
#delta_ij <- matrix(runif(n^2, min = 0.02, max = 0.05), ncol = n)
d_ij <- matrix(rep(0,n^2), ncol = n) # turn off diffusion terms
diag(d_ij) <- 0
A_ij <- matrix(rbinom(n^2, 1, prob = 0.3), n, n)
diag(A_ij) <- 0
## Parameters: use rep() if the value is the same, or runif() ir meant to be different across systems
params <- list(
    r = rep(1.5, n),              # intrinsic growth rate
    K = rep(10, n),               # Carrying capacity
    c = rep(2, n),                # max predation rate
    q = runif(n, min = 2, max = 8),        # threshold
    b = 4 ,                       # sharpness of the shift
    d_ij = d_ij,                  # matrix of difussion terms
    A_ij = A_ij                   # adjacency matrix
)

## set up time steps
times <- seq(from = 0, to = 100, by = 0.01)

## initial conditions
yini <- runif(n, 5, 20)

## run the model
#print(system.time(
    out <- ode(
      y = yini, times = times,  func = resource, parms = params,
      #method = "bdf" , ## see help("ode") for more methods
      events=list(func = posfun, time = times)
    )
#))

```
Model with no connections:

```{r warning = FALSE, message = FALSE}
df <- out %>% as_tibble() %>%
    gather(key = "species", value = "population", 2:last_col())

g1 <- df %>%
    ggplot(aes(x=time, y=population)) +
    geom_line(aes(color = species), size = 0.25, show.legend = T) + 
    theme_light() + ylim(c(0,20)) +
    labs(title = "No connections", tag = "A")
```

Connected model: 

```{r warning = FALSE, message = FALSE}
## set up flux matrix
d_ij <- matrix(runif(n^2, min = 0.02, max = 0.05), ncol = n)
diag(d_ij) <- 0

## Only change the delta:
params$d_ij <-  d_ij           # adjacency matrix of difussion terms

## run the model
out <- ode(
  y = yini, times = times,  func = resource, parms = params,
  method = "bdf", ## see help("ode") for more methods
  events=list(func = posfun, time = times)
)


df <- out %>% as_tibble() %>%
    gather(key = "species", value = "population", 2:last_col())

g2 <- df %>%
    ggplot(aes(x=time, y=population)) +
    geom_line(aes(color = species), size = 0.25, show.legend = T) + 
    theme_light() + ylim(c(0,20)) +
    labs(title = "Connected", tag = "B")

## plot the network

net <- network::network(A_ij, directed = T)

p1 <- ggnetwork(net) %>%
  ggplot() + 
  geom_edges(
    aes(x = x,y = y,xend = xend,yend = yend), color = "gray50", size = 0.1,
    arrow = arrow(length = unit(0.15, "lines"), type = "closed"),
    curvature = 0.1) +
  geom_nodes(aes(x =x,y = y, color = as.factor(vertex.names)), 
             size = 2, show.legend = FALSE ) +
  labs(tag= "C") +
  theme_void()

g1 + g2 + p1 + plot_layout(guides = "collect")

```



## Normalized model:

$$  
\frac{dY_i}{d\tau} =R_iY_i\left(1-\frac{Y_i}{K_i}\right)-C_i \frac{Y_i^{\beta_i}}{1+Y_i^{\beta_i}} + \sum_{j\not=i}D_{ij}(Q_{ij}Y_j-Y_i)
$$

Let $r$ denote the time difference between $t$ and $\tau$, $R_i=\frac{r_i}{r}$, $K_i=\frac{k_i}{q_i}$, $C_i=\frac{c_i}{rq_i}$, $Q_{ij}=\frac{d_{ji}q_j}{d_{ij}q_i}$, and $D_{ij}=\frac{d_{ij}}{r}$. 

```{r warning = FALSE, message = FALSE}
## Pollution model
resource_norm <- function(t, y, params){
  with(as.list(c(y, params)), {
    ## Normalization steps:
    Y = y/r0
    R = r/r0
    K0 = K/q
    C = c/(r0*q)
    Qij_parts <- d_ij * q
    Qij = t(Qij_parts) / Qij_parts ## the same as lower triangle / upper triangle and viceversa
    ## the next line corrects the case when Zij is NaN induced by division by zero. In that case, Zij should be a matrix full of zeroes = no connections.
    if(all(is.nan(Qij))) Qij = Qij_parts
    
    # set diag to zero
    diag(Qij) <- 0
    D_ij = d_ij / r0
    
    ## System with normalized parameters
    resource <- R * Y * (1 - (Y/K0)) - C * (Y^b / (1 + Y^b))
    outflow <-  (A_ij * D_ij * Qij) %*% Y
    inflow <-  t(A_ij * D_ij) %*% Y 
    # Qij is not multiplied by D_ij by suggestion of ASC.

    dX <- resource + (inflow - outflow)
    
    return(list(c(dX)))
  })
}

# to make it comparable with the non-normalized version I need to keep the 
# same parameters. Only update the scaling parameter
# params <- list(
#     r0 = rep(1,n),                # scaling parameter
#     r = rep(1.5, n),              # intrinsic growth rate
#     K = rep(10, n),               # Carrying capacity
#     c = rep(2, n),                # max predation rate
#     q = runif(n, min = 2, max = 8),        # threshold
#     b = 4 ,                       # sharpness of the shift
#     d_ij = d_ij,                  # matrix of diffusion terms
#     A_ij = A_ij                   # adjacency matrix
# )
params$r0 <- rep(1,n)

## run the model
out <- ode(
  y = yini, times = times,  func = resource_norm, parms = params,
  method = "bdf", ## see help("ode") for more methods
  events=list(func = posfun, time = times)
)


df <- out %>% as_tibble() %>%
    gather(key = "species", value = "population", 2:last_col())

g3 <- df %>%
    ggplot(aes(x=time, y=population)) +
    geom_line(aes(color = species), size = 0.25, show.legend = T) + 
    ylim(c(0,20)) +
    theme_light() + labs(tag = "A", title = "Normalized")

g3 + g2 + p1 + plot_layout(guides = "collect")
```

## To-do:

1. Systems without connections: Now with a working model, update and change some parameters.
```{r warning=FALSE, message=FALSE}
# delta_ij # old delta_ij used in previous runs
## Turn-off the connection
#params$delta_ij <- matrix(rep(0, n^2), n, n)
params$A_ij <- matrix(rep(0, n^2), n, n)

## run the model
out <- ode(
  y = yini, times = times,  func = resource_norm, parms = params,
  #method = "bdf" ## see help("ode") for more methods
  events=list(func = posfun, time = times)
)


g4 <- out %>% as_tibble() %>%
    gather(key = "system", value = "pollutant", 2:last_col()) %>%
    ggplot(aes(x=time, y=pollutant)) +
    geom_line(aes(color = system), size = 0.25, show.legend = T) + 
    ylim(c(0,20)) +
    theme_light() + labs(tag = "B", title = "no connections")

g3 + g4 + p1+ plot_layout(guides = "collect")

```



2. Systems with connections (max 10)

I first update N to 10 and create initial conditions for all systems. For comparison I run a model without connection, and then turn on diffusion. 

```{r warning=FALSE, message=FALSE}
## I need to update all parameters then:
## set up flux matrix
n <- 10
#delta_ij <- matrix(runif(n^2, min = 0.02, max = 0.05), ncol = n)
d_ij <- matrix(rep(0,n^2), ncol = n) # turn off diffusion
diag(d_ij) <- 0
A_ij <-  matrix(rep(0,n^2), ncol = n)
diag(A_ij) <- 0
## Parameters: use rep() if the value is the same, or runif() if meant to be different across systems
params <- list(
    r0 = rep(1,n),                # scaling parameter
    r = rep(1.5, n),              # intrinsic growth rate
    K = rep(10, n),               # Carrying capacity
    c = rep(2, n),                # max predation rate
    q = runif(n, min = 2, max = 8),        # threshold
    b = 4 ,                       # sharpness of the shift
    d_ij = d_ij,                  # matrix of diffusion terms
    A_ij = A_ij                   # adjacency matrix
)

## initial conditions
yini <- runif(n, 5, 20)
## run the model
out <- ode(
  y = yini, times = times,  func = resource_norm, parms = params,
  #method = "bdf" ## see help("ode") for more methods
  events=list(func = posfun, time = times)
)

## Visualize
g5 <- out %>% as_tibble() %>%
    gather(key = "system", value = "resource", 2:last_col()) %>%
    ggplot(aes(x=time, y=resource)) +
    geom_line(aes(color = system), size = 0.25, show.legend = T) + 
    theme_light() + labs(tag = "A", title = "Disconnected")
```

Note that currently the $d_{ij}$ matrix has low diffusion coefficients (2-5%) but all lakes are connected to each other. To better control structure, one should multiply the difussion coefficents to a [0,1] adjacency matrix, so one gets control of structure on a term by itself. [feature added]

```{r warning = FALSE, message=FALSE}
## get a erdos-renyi random graph
A_ij <- igraph::sample_gnp(n, 0.1, TRUE) %>% 
  igraph::as_adjacency_matrix(graph = .) %>% 
  as.matrix()

## activate conections
d_ij <- matrix(runif(n^2, min = 0.02, max = 0.05), ncol = n)
diag(d_ij) <- 0
params$d_ij <- d_ij
params$A_ij <- A_ij

## run the model
out <- ode(
  y = yini, times = times,  func = resource_norm, parms = params,
  #method = "bdf" ## see help("ode") for more methods
  events=list(func = posfun, time = times)
)

# Visualize
g6 <- out %>% as_tibble() %>%
    gather(key = "system", value = "resource", 2:last_col()) %>%
    ggplot(aes(x=time, y=resource)) +
    geom_line(aes(color = system), size = 0.25, show.legend = T) + 
    theme_light() + labs(tag = "B", title = "Connected")

## network
net <- network::network(A_ij, directed = T)

p1 <- ggnetwork(net) %>%
  ggplot() + 
  geom_edges(
    aes(x = x,y = y,xend = xend,yend = yend), color = "gray50", size = 0.1,
    arrow = arrow(length = unit(0.15, "lines"), type = "closed"),
    curvature = 0.1) +
  geom_nodes(aes(x =x,y = y, color = as.factor(vertex.names)), size = 2, show.legend = FALSE ) +
  theme_void(base_size = 4)
  


g5 + g6 + p1 + plot_layout(guides = "collect")
```


3. Scenario when Q_ij > 1 vs. Q_ij < 1

Because $Q_{ij}=\frac{d_{ji}z_j}{d_{ij}z_i}$ and $q_i = q_j$, the only way to control $Q$ in a way that it becomes >1 or <1 depends on $d_{i,j}$. The matrix needs to be created with the upper triangle > lower triangle or viceversa.

```{r warning = FALSE, message=FALSE}
## This makes sure that Z_ij > 1
d_ij <- matrix(0, ncol = n, nrow = n)
d_ij[upper.tri(d_ij) ] <-  0.02
d_ij[lower.tri(d_ij) ]<- 0.05

params$d_ij <- d_ij

## run the model
out <- ode(
  y = yini, times = times,  func = resource_norm, parms = params,
  #method = "bdf" ## see help("ode") for more methods
  events=list(func = posfun, time = times)
)

## This makes sure that Z_ij > 1
d_ij <- matrix(0, ncol = n, nrow = n)
d_ij[upper.tri(d_ij) ] <-  0.05
d_ij[lower.tri(d_ij) ]<- 0.02

params$d_ij <- d_ij

## run the model
out2 <- ode(
  y = yini, times = times,  func = resource_norm, parms = params,
  #method = "bdf" ## see help("ode") for more methods
  events=list(func = posfun, time = times)
)

out <- out %>% as_tibble() %>%
  gather(key = "system", value = "pollutant", 2:last_col()) %>%
  add_column(scenario = "Z > 1") %>%
  mutate(time = as.numeric(time), pollutant = as.numeric(pollutant))
  
out2 <- out2 %>% as_tibble() %>%
  gather(key = "system", value = "pollutant", 2:last_col()) %>%
  add_column(scenario = "Z < 1") %>% 
  mutate(time = as.numeric(time), pollutant = as.numeric(pollutant))

# Visualize
p2 <- full_join(out,out2) %>% 
    ggplot(aes(x=time, y=pollutant)) +
    geom_line(aes(color = system), size = 0.25, show.legend = FALSE) + 
  facet_wrap(~scenario) +
    theme_light(base_size = 6)
```

```{r}
library(patchwork)

p2 + p1 + plot_layout(width = c(2,1)) 

# ggsave("figure_Z_comparison.png", width = 5, height = 2, units = "in", dpi = 800)
```






However, not all $Q_{ij}$ are >1 or < 1. 

```{r}
qmat <- with(params, d_ij * q)
(t(qmat)/qmat) > 1
```

The reason for that is that it depends also of $q_i$ on the original equation. The only way to control for that is to constructing $q$ by hand, or making it the same for all systems.

4. Scenario when $D_ij > 1$ versus $D_{ij} < 1$

```{r warning = FALSE, message=FALSE}
## reset d_ij
d_ij <- matrix(runif(n^2, min = 0.02, max = 0.05), ncol = n)
diag(d_ij) <- 0
params$d_ij <- d_ij

## By setting s to > min value of delta, the matrix is >=1 always
params$s <- 0.015

## run the model
out <- ode(
  y = yini, times = times,  func = resource_norm, parms = params,
  method = "bdf" ,## see help("ode") for more methods
  events=list(func = posfun, time = times)
)

## This makes sure that Z_ij > 1
## By setting s to > max value of delta, the matrix is <= 1 always
params$s <- 0.07

## run the model
out2 <- ode(
  y = yini, times = times,  func = resource_norm, parms = params,
  method = "bdf", ## see help("ode") for more methods
  events=list(func = posfun, time = times)
)

out <- out %>% as_tibble() %>%
  gather(key = "system", value = "pollutant", 2:last_col()) %>%
  add_column(scenario = "Delta_{ij} > 1") %>%
  mutate(time = as.numeric(time), pollutant = as.numeric(pollutant))
  
out2 <- out2 %>% as_tibble() %>%
  gather(key = "system", value = "pollutant", 2:last_col()) %>%
  add_column(scenario = "Delta_{ij} < 1") %>% 
  mutate(time = as.numeric(time), pollutant = as.numeric(pollutant))

# Visualize
full_join(out,out2) %>% 
    ggplot(aes(x=time, y=pollutant)) +
    geom_line(aes(color = system), size = 0.25, show.legend = T) + 
  facet_wrap(~scenario) +
    theme_light()
```

5. Test for structural stability before running simulations
6. ~~Repeat for resource system. [done]~~
7. Set up controlling experiments

Notes after meeting with ASC: 

* ~~fix the $A_{ij}$, the model was not updated [done]~~
* ~~extract results and create a graph with scenarios when Z and Delta are >1<~~
* phase diagrams X vs U?

