---
title: "Simple model of interaction between resource and pollution systems"
output:
  html_document:
    df_print: paged
    code_folding: hide
    toc: true
    toc_float:
      collapsed: true
      smooth_control: true
    toc_depth: 3
    fig_caption: true
    fig_width: 4
    fig_height: 3.5
    highlight: tango
    self_contained: true
    theme: "paper"
  html_notebook:
    df_print: paged
    code_folding: hide
    toc: true
    toc_float:
      collapsed: true
      smooth_control: true
    toc_depth: 3
    fig_caption: true
    fig_width: 4
    fig_height: 3.5
    highlight: tango
    self_contained: true
    theme: "paper"
  #pdf_document: default
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(comment=NA, fig.width=4, fig.height=3.5, tidy = TRUE)
```


## Simulations


ASC did some simulations in Matematica that I here try to replicate in `R`.
First I load some libraries necesary for solving the model numerically and plotting:

```{r libraries, warning=FALSE, message=FALSE}
library(deSolve)
# library(deTestSet)
library(tidyverse)
library(patchwork)
```

### Normalized model

The replicated model is: 

$$
\frac{dX_i}{d\tau} = R_i X_i \left( 1 - \frac{X_i}{K_i}\right) -
C_i\frac{(X_i)^{a_i}}{1 + (X_i)^{a_i}} +
\sum_{j \neq i}D_{ij}(Q_j X_j - X_i)
$$

where the initial conditions are:

```{r initial_conditions}

n <- 2 ## number of cresources

## Parameters: The delta and d parameters are set here by hand to replicate ASC results.
##  Later they are set randomly on the fly.
R <- 1          # Growth rate
K <- 10         # carryng capacity
delta <- 0.05   # share of stock that diffuses to other plots
a <- 4          # maximum level after crossing the threshold
c <- 1.7        # consumption efficiency
d <- 0.02       # total diffiusion from all other plots to i
Dij <- matrix(c(0, delta, d, 0), nrow = 2, byrow = FALSE)

## Initial conditions:

yini <- runif(n, 8, 20)  

```

```{r, fig.align="center", fig.width=1, fig.height=1}
## small dataframe
df <- tibble( x = rep(0:12, 3))
## components of the function, with and without diffusion
growth <- function(x) {R*x*(1-(x/K))}
growth_diffusion <- function(x) {R*x*(1-(x/K)) + d*x - delta *x}
consumption <- function(x){c *((x^a) / (1 + (x^a) ))}
## plot
ggplot(data = df, aes(x)) +
  stat_function(fun = growth, color = "blue", size = 0.1) +
  stat_function(fun = growth_diffusion, color = "orange", size = 0.1) +
  stat_function(fun = consumption, color = "purple", size = 0.1) +
  geom_hline(yintercept = 0, color = "gray", size = 0.1) + 
  labs(x = "x", y = "f(x)") + #xlim(0,12) + ylim(0,3) +
  theme_light(base_size = 7)

```

then we specify the model with two stocks, so we can make use of the difference between `delta` and `d` parameters:

```{r}
## The problem with the function I had is that the response variable needs to be 
## named y inside the function. 
resource <- function(t, y, params){
    with(as.list(y, params), {
        growth <- R * y * (1- y/K)  ## This should recover Lodka-Volterra
        consumption <- c *((y^a) / (1 + (y^a) ))
        
        immigration <-  t(Dij) %*% y
        emmigration <-  Dij %*% y
    
        dX <- growth - consumption + (immigration - emmigration)

    return(list(c(dX)))
    })
}
```
Now ready to run the model:

```{r}
## set up time steps
times <- seq(from = 0, to = 100, by = 0.01)
## declare parameters
params <- list(a = a, R = R, K = K, c = c, Dij = Dij)

## run the model
print(system.time(
    out <- ode(
      y = yini, times = times, parms = params, func = resource,
      method = "bdf" ## see help("ode") for more methods
    )
))

```

```{r warning=FALSE, message=FALSE}
n <- 2
df <- out %>% as_tibble() %>%
    gather(key = "species", value = "population", 2:last_col())

df %>%
    ggplot(aes(x=time, y=population)) +
    geom_line(aes(color = species), size = 0.25, show.legend = T) + 
    ylim(c(0,20)) +
    theme_light()
```

Now as expected, the two species reached carrying capacity over time.

### Resource model with many species

```{r warning=FALSE, message=FALSE}
## Initial conditions:
n <- 5 ## number of resources
yini <- runif(n, 10, 20)  
## heterogeneous case
Dij <- matrix(runif(n^2, min = 0, max = 0.025), ncol = n) # model very sensitive to diffusion
diag(Dij) <- 0

## Parameters
params <- list(
  R = 1,        # Growth rate
  K = 10,       # carryng capacity
  a = 4,         # maximum level after crossing the threshold
  c = 1.7,
  Dij = Dij
)


## run the model
print(system.time(
    out <- ode(
      y = yini, times = times, parms = params, func = resource,
      method = "bdf" ## see help("ode") for more methods
    )
))

## J200218: method "bdf" is good for stiff-problems, see Soetaert 2012 p
```

```{r warning=FALSE, message=FALSE}
out %>%
  as_tibble() %>% 
  gather(key = "species", value = "abundance", 2:6) %>%
  ggplot(aes(time, abundance)) + 
  geom_line(aes(color = species)) +
  theme_light()
```


#### Some ideas:
- **Can you modify the model with a ifelse statement to avoid negative values?**: The `ifelse` option did not work as expected. I found the following solution in [Stackoverflow](https://stackoverflow.com/questions/49883304/ordinary-differential-equations-odes-is-there-any-way-to-prevent-negative-va/49933276#49933276) using an additional event function.

```{r}

# This is where we define your event function
# Add this directly above your call to ode()
  posfun <- function(t, y, parms){
    with(as.list(y), {
      y[which(y<0)] <- 0  
      return(y)
    })
  }

## the function is the same as above: resource

out <- ode(y = yini, 
           times = times, 
           parms = params, 
           func = resource,
           method = "bdf",
           events=list(func = posfun, time = times))
```

- **On the resource only model, how do you induce regime shifts?** slow variables, noise? which variables make more sense?
This first example below changes the consumption rate or $c$ in the equation for the resource model.

```{r}
## slow variable 
# varying consumption term

df_consumption <- data.frame(year = seq(0, 100, 1), c = seq(0.5, 5, length.out = 101 ))
c_var <- approxfun(df_consumption)

## Parameters
params <- list(
  R = 0.5,        # Growth rate
  K = 10,       # carryng capacity
  a = 4,         # maximum level after crossing the threshold
  # c = 1.7,    ## This is the value we are changing over time
  Dij = Dij
)

resource_c_var <- function(t, y, params){
    with(as.list(y, params), {
      
      c <- c_var(t)
      
        growth <- R * y * (1- y/K)  ## This should recover Lodka-Volterra
        consumption <- c *((y^a) / (1 + (y^a) ))
        
        immigration <-  t(Dij) %*% y
        emmigration <-  Dij %*% y
    
        dX <- growth - consumption + (immigration - emmigration)
        
        ## this is the ifelse try that didn't work.
        # dX <- ifelse(
        #   growth - consumption + (immigration - emmigration) <= 0, 0,
        #   growth - consumption + (immigration - emmigration)
        # ) 
        
    return(list(c(dX), c = c))
    })
}

## run the model
print(system.time(
    out <- ode(
      y = yini, times = times, parms = params, func = resource_c_var,
      method = "bdf", ## see help("ode") for more methods
      events=list(func = posfun, time = times)
    )
))

```


```{r warning=FALSE, message=FALSE}
out %>%
  as_tibble() %>% 
  gather(key = "species", value = "abundance", 2:last_col()) %>%
  ggplot(aes(time, abundance)) + 
  geom_line(aes(color = species)) +
  theme_light()
```

J200219:Next steps with only resource model is to do a sensitivity analysis of the coupling parameter $D_ij$ when $d_i=d_j$ and $d_i > d_j | d_j > d_i$. I can play with directionality on the adjacency matrix to see the differences on direction of flow as well. The expectation is a delay or earlier collapse depending of domino flow. It would be cool to visualize the network as well with link and coupling strenght and use coordinates to show the expectation of who collapses first on the domino and who's next.



### Pollution model:

**J200225:** Need to modify the models so it works on one dimension (one resource or one lake alone) as well.

$$\forall{i} \in\{1,...,n\}; \frac{dy_i}{dt} = u_i - b_iy_i + v_i \frac{y_{i}^{\beta_{i}}}{Y^{\beta_{i}} + y_{i}^{\beta_{i}}} 
+ \sum_{j \neq i} \Delta_{ij} (y_{j} - y_{i})$$

```{r}

# This is where we define your event function
# Add this directly above your call to ode()
posfun <- function(t, y, parms){
  with(as.list(y), {
    y[which(y<0)] <- 0
    return(y)
  })
}

## Pollution model:

pollution <- function(t, y, params){
  with(as.list(c(y, params)), { ## adding the c() eliminates the error of not finding object 'u'
    
    pollutant <- u - (b*y) + v * (y^beta/(Y^beta + y^beta))
    outflow <-  Dij %*% y
    inflow <-  t(Dij) %*% y

    dy <- pollutant + (inflow - outflow)
    
    return(list(c(dy)))
  })
}

## set up flux matrix
n <- 3
Dij <- matrix(
  #runif(n^2, min = 0.02, max = 0.05), 
  0,
  ncol = n, nrow = n)
diag(Dij) <- 0

## Parameters
#params <- list(u = 0.5, v = 1.7, Y = 1, beta = 4, Dij = Dij )
params <- list(
  b = 2.2,
  v = 10,
  u = 0.5,
  Y = 2.5,
  beta = 4 ,
  Dij = Dij
)


yini <- runif(n, 0, 5)

## run the model
print(system.time(
    out <- ode(
      y = yini, times = times, parms = params, func = pollution,
      method = "bdf", ## see help("ode") for more methods
      events=list(func = posfun, time = times)
    )
))

```
Simulation results:
```{r warning=FALSE, message=FALSE}
out %>%
  as_tibble() %>% 
  gather(key = "lakes", value = "pollutants", 2:4) %>%
  ggplot(aes(time, pollutants)) + 
  geom_line(aes(color = lakes)) + 
  ylim(0,20)+
  theme_light()
```

So indeed, depending on initial conditions now we have lakes that go to almost no pollutants in the water (oligotrophic), and lakes with "high" level of pollutants (eutrophic). What I don't get now is why the equilibrium on the high nutrient regime is 4.31 when the figure below shows its intersection around  a value of 8. Note that the parameters are the same.

This model has already the in-built function that prevents the behaviour of having negative pollutants.

```{r warning=FALSE, message=FALSE}
## parameters on the working environment
  v = 10    # this is like a carrying capacity
  u = 0.5
  Y = 2.5
  beta = 4
  Dij = Dij
  b = 2.2


## small dataframe
df <- tibble( x = rep(0:12, 3))
## components of the function, with and without diffusion
pollutant <- function(y) { v * (y^beta/(Y^beta + y^beta))}  ## Original model formulated by ASC
pollutant_sedimentation <- function(y) { v * (y^beta/(Y^beta + y^beta)) } ## adding the saturation parameter
management <- function(y){- u + (b*y)} ## now the diffusion term is in the management line.
management_diffusion <- function(y){ - ((0.01 * y) - (0.05 * y)) - u + (b*y)} ## now the diffusion term is in the management line.
## plot
ggplot(data = df, aes(x)) +
  #stat_function(fun = pollutant, color = "blue", linetype = 2) +
  stat_function(fun = pollutant_sedimentation, color = "orange") +
  stat_function(fun = management_diffusion, color = "purple") +
  stat_function(fun = management, color = "dodgerblue") +
  geom_hline(yintercept = 0, color = "gray") + 
  labs(x = "y", y = "f(y)") +
  theme_light()
```

Cause a regime shift in the lake system with and without diffusion:

```{r warning = FALSE, message = FALSE}
## slow variable 
# varying consumption term

df_pollution <- data.frame(year = seq(0, 100, 1), c = seq(0.25, 5, length.out = 101 ))
u_var <- approxfun(df_pollution)

## Pollution model:

pollution_var <- function(t, y, params){
  with(as.list(c(y, params)), { ## adding the c() eliminates the error of not finding object 'u'
    
    u <- u_var(t)
    
    pollutant <- u - (b*y) + v * (y^beta/(Y^beta + y^beta))
    outflow <-  Dij %*% y
    inflow <-  t(Dij) %*% y

    dy <- pollutant + (inflow - outflow)
    
    return(list(c(dy), u = u))
  })
}

## set up flux matrix without diffusion
n <- 5
Dij <- matrix(
  #runif(n^2, min = 0.02, max = 0.05), 
  0,
  ncol = n, nrow = n)
diag(Dij) <- 0


## Parameters
params <- list(
  b = 2.2,
  v = 10,
 # u = 0.5,
  Y = 2.5,
  beta = 4 ,
  Dij = Dij
)

yini <- runif(n, 0, 5)

## run the model
print(system.time(
    out <- ode(
      y = yini, times = times, parms = params, func = pollution_var,
      method = "bdf", ## see help("ode") for more methods
      events=list(func = posfun, time = times)
    )
))

out %>%
  as_tibble() %>% 
  gather(key = "lakes", value = "pollutants", starts_with("1"):last_col()) %>%
  ggplot(aes(time, pollutants)) + 
  geom_line(aes(color = lakes)) + 
  ylim(0,20)+
  theme_light()


```

Now with diffusion, same parameters:

```{r warning = FALSE, message=FALSE}
## set up flux matrix without diffusion
n <- 5
Dij <- matrix(
  runif(n^2, min = 0.02, max = 0.05), 
  #0,
  ncol = n, nrow = n)
diag(Dij) <- 0


## Parameters
params <- list(
  b = 2.2,
  v = 10,
  # u = 0.5,
  Y = 2.5,
  beta = 4 ,
  Dij = Dij
)

## run the model
print(system.time(
    out <- ode(
      y = yini, times = times, parms = params, func = pollution_var,
      method = "bdf", ## see help("ode") for more methods
      events=list(func = posfun, time = times)
    )
))

out %>%
  as_tibble() %>% 
  gather(key = "lakes", value = "pollutants", starts_with("1"):last_col()) %>%
  ggplot(aes(time, pollutants)) + 
  geom_line(aes(color = lakes)) + 
  ylim(0,20)+
  theme_light()
```

On the lake only model with diffusion, because some lakes can remain on the clear state a bit longer given the extra load of nutrients lost through diffusion. Because the same reason, lakes that were oligotrophic can tip to the eutrophic state later on. So diffusion can delay the onset of the regime shift if the system is outsorcing nutrients, or make the shift earlier if the lake is taking extra nutrients from another lake. The change of timing depends on network structure.

Up until now the diffusion term has been a random matrix that generally couple all lakes with each other on a dense network with very weak coupling. So the dynamics are dominated by what is going on locally, but diffusion can play an important role. Here is an example of one of the random matrixes that is currently on our last model $D_{ij}$

```{r}
Dij %>%
  as_tibble() %>%
  rownames_to_column(var = "from") %>%
  pivot_longer(cols = V1:last_col(), names_to = "to",
               values_to = "delta"
               ) %>%
  mutate(to = str_remove_all(to, "V") ) %>%
  ggplot(aes(to, from)) +
  geom_tile(aes(fill = delta)) +
  theme_minimal()
```
In discussion with Anne-Sophie we decided to have a tree structure on the network for the pollution model. The intuition is that lakes connected by water streams usually do not form loops due to gravity. To implement that on the model, first I use the `igraph` package to produce tree-networks, or directed acyclic graphs (DAGs). Next, I introduce two different terms on the pollution system equation: $A_{ij}$ is the adjacency matrix of the network whose values are either 0,1; $D_{ij}$ is the diffusion parameter. If it's a scalar, the diffusion is the same for all lakes. If it's a random matrix, it will asign random diffusion coefficients to each lake connection in $A_{ij}$. If it's a vector, one can play with the ordering of the vector to study the cases of $d_i > d_j | d_j > d_i$. For now I will simulate 10 lakes with the following network structure:

```{r}

igraph::make_tree(n = 10, mode = "out", children = 2) %>% igraph::plot.igraph(layout = igraph::layout_as_tree)
 # igraph::as_adj(sparse = FALSE)
```

The expectation with this particular tree structure is: 

* Lake 1 will be delayed du to lack of input
* Lakes 2-3 will be delayed because $d_{in} = 0.025 > d_{out} = 0.05$
* Lakes 4-5 will be delayed because $d_{in} > d_{out}$
* Lakes 6-7 will be faster because $d_{in} > d_{out} = 0$
* Lakes 8-9 will be faster because $d_{in} > d_{out} = 0$
* Lakes 10 faster than 8-9, $d_{in}$ is higher.

```{r warning = FALSE, message=FALSE, fig.width=5, fig.height=3.5}
## set up flux matrix without diffusion
n <- 10
Aij <- igraph::make_tree(n = n, mode = "out", children = 2) %>% 
  igraph::as_adj(., sparse = FALSE)

Dij <- matrix(
  runif(n^2, min = 0.02, max = 0.05), 
  ncol = n, nrow = n)
diag(Dij) <- 0

Dij <-  Dij * Aij # pair-wise multiplication for d_i != d_j
Dij <-  Aij * 0.05 # scalar multiplication for d_i == d_j

## Parameters
params <- list(
  b = 2.2,
  v = 10,
  # u = 0.5,
  Y = 2.5,
  beta = 4 ,
  Dij = Dij
)

yini <- rep(0.5, n) # same initial conditions to all
  
  
## run the model
print(system.time(
    out <- ode(
      y = yini, times = times, parms = params, func = pollution_var,
      method = "bdf", ## see help("ode") for more methods
      events=list(func = posfun, time = times)
    )
))

p <- out %>%
  as_tibble() %>% 
  gather(key = "lakes", value = "pollutants", starts_with("1"):last_col()) %>%
  ggplot(aes(time, pollutants)) + 
  geom_line(aes(color = lakes)) + 
  ylim(0,10)+
  theme_light()

plotly::ggplotly(p)
```

For comparison, the same model without any connections:

```{r warning = FALSE, message=FALSE}
Dij <- matrix(0, nrow = n, ncol = n)

params <- list(
  b = 2.2,
  v = 10,
  # u = 0.5,
  Y = 2.5,
  beta = 4 ,
  Dij = Dij
)

out <- ode(
      y = yini, times = times, parms = params, func = pollution_var,
      #method = "bdf", ## see help("ode") for more methods
      events=list(func = posfun, time = times))

g2 <- out %>%
  as_tibble() %>% 
  gather(key = "lakes", value = "pollutants", starts_with("1"):last_col()) %>%
  ggplot(aes(time, pollutants)) + 
  geom_line(aes(color = lakes)) + 
  ylim(0,10)+
  theme_light()

plotly::ggplotly(g2)
```



With constant $d_i = d_j$, the lakes downstream shift first, partly because they don't have outflow and accumulate faster the nutrients with respect to their sedimentation terms.  

Some ideas for future analysis:

- Record the time at which the system flips. How does the timing change depending on (i) system syze, and (ii) network structure? => random network, tree-network, network with feedbacks, scale-free(?), degree / density.
- Right now the rate of pollution $u$ is affecting all lakes at the same time, either as a constant on the simple model, or as a slow variable in the latest example. What if one only tips one lake? -- I will need a $u$ function per lake, possible but how do I code it?

